---
title: "Homework 8"
author: "Benjamin Noland"
output: pdf_document
fontsize: 12pt
---

```{r setup, include = FALSE}
library(TSA)

#options(digits = 3)

knitr::opts_chunk$set(
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  fig.show = "hold",
  eval = TRUE
)
```

\newcommand{\var}{\mathrm{Var}}
\newcommand{\cov}{\mathrm{Cov}}
\newcommand{\corr}{\mathrm{Corr}}
\newcommand{\E}{\mathrm{E}}
\newcommand{\pr}{\mathrm{P}}

1.
2. (Cryer & Chan, Exercise 10.9)
    a. A plot of the original time series is shown below:
    ```{r}
    data(airpass)
    plot(airpass, type = "o")
    ```
    This plot shows that the time series is clearly non-stationary. In
    particular, it has non-constant mean, and the series grows more variable
    with time. After taking logarithms, we get the following:
    ```{r}
    plot(log(airpass), type = "o")
    ```
    Although this transformed series appears to still have non-constant mean,
    the series no longer appears to be getting more variable with time. Thus the
    transformation seems appropriate.
    
    b. After taking first differences of the log-transformed series, we get the
    following:
    ```{r}
    first_diff <- diff(log(airpass))
    plot(first_diff, type = "o")
    ```
    This new series appears to be stationary (in particular, it has taken care
    of the issue of non-constant mean in the undifferenced series). However,
    there still appears to be seasonality in the series.
    
    c. It seems reasonable to expect a seasonality of $s = 12$. Taking seasonal
    differences with $s = 12$, we get the following:
    ```{r}
    seasonal_diff <- diff(first_diff, lag = 12)
    plot(seasonal_diff, type = "o")
    ```
    The seasonal differencing appears to have removed the seasonality in the
    series.
    
    d. A plot of the sample ACF of the seasonal difference of the first
    difference of the log-transformed series is as follows:
    ```{r}
    acf(as.vector(seasonal_diff))
    ```
    The ACF is statistically significant from zero at lag 1, and then (with a
    few exceptions), is zero for higher lags. This suggests that this
    transformed series *may* be suitably modeled by an MA(1) process.
    
    e. The following fits an ARIMA$(0, 1, 1) \times (0, 1, 1)_{12}$ model to the
    log-transformed series:
    ```{r}
    fit <- arima(log(airpass), order = c(0, 1, 1),
                 seasonal = list(order = c(0, 1, 1)))
    fit
    ```

    f.

3.
4. (Cryer & Chan, Exercise 13.13)
5. (Cryer & Chan, Exercise 13.17)
6. (Cryer & Chan, Exercise 13.21)
7. (Cryer & Chan, Exercise 12.9)
